---
title: "P8105 Homework 1"
author: "Junxian Chen (jc5314)"
date: "09-14-2019"
output: github_document
---

# Problem 1

### 1. Create a data frame called `p1_data`:

```{r, message = FALSE}
library(tidyverse)

# data frame for problem 1
p1_df = tibble(                  
  norm_samp = rnorm(8),
  norm_samp_pos = norm_samp > 0,
  character_vector = LETTERS[1:8],
  factor_level = gl(3, 3, length = 8, labels = c("X", "Y", "Z"))
)
```

### 2. Take the mean of each variable in the data frame `p1_data`:

```{r, }
mean(pull(p1_df, norm_samp))
mean(pull(p1_df, norm_samp_pos))
mean(pull(p1_df, character_vector))
mean(pull(p1_df, factor_level))
```

When trying to take the mean of each variable in the data frame, the first two variables work and give numeric answers. While the third and forth variable do not work and give out warning messages and `NA` as result. 

### 3. Convert variables from one type to another:

Applies the `as.numeric` function to the logical, character, and factor variables:

```{r, eval = FALSE}
as.numeric(pull(p1_df, norm_samp_pos))
as.numeric(pull(p1_df, character_vector))
as.numeric(pull(p1_df, factor_level))
```

**(What happens, and why? Does this help explain what happens when you try to take the mean?)**

```{r, eval = FALSE}
# Convert the logical vector to numeric, and multiply the random sample by the result:

as.numeric(pull(p1_df, norm_samp_pos)) * pull(p1_df, norm_samp)

# Convert the logical vector to a factor, and multiply the random sample by the result:

factor(pull(p1_df, norm_samp_pos)) * pull(p1_df, norm_samp)

# Convert the logical vector to a factor and then convert the result to numeric, and 
# multiply the random sample by the result:

factor_samp_pos = factor(pull(p1_df, norm_samp_pos))
as.numeric(factor_samp_pos) * pull(p1_df, norm_samp)
```

# Problem 2

### 1. Creat a data frame called `p2_df`:

```{r}
# data frame for problem 2
p2_df = tibble(                  
  x = rnorm(500),
  y = rnorm(500),
  xy_sum_compare = x + y > 1,
  xy_sum_compare_numeric = as.numeric(xy_sum_compare),
  xy_sum_compare_factor = factor(xy_sum_compare)
)
```

Dataset description: 

The dataset `p2_df`contains `r nrow(p2_df)` rows and `r ncol(p2_df)` columns. For the variable `x` in dataset, the mean is `r mean(pull(p2_df, x))`, the median is `r median(pull(p2_df, x))` and the standard deviation is `r sd(pull(p2_df, x))`. The proportion of cases for which `x + y > 1` is `r sum(pull(p2_df, xy_sum_compare)) / length(pull(p2_df,x)) * 100`%.

